### 2016-10-07 mu/mpxerr.tex
up Dab, ~Dab, and -Dab plots.

\documentclass{article}
\begin{document}

\gdef\mpxshipout{\shipout\hbox\bgroup
  \setbox0=\hbox\bgroup}
\gdef\stopmpxshipout{\egroup  \dimen0=\ht0 \advance\dimen0\dp0
  \dimen1=\ht0 \dimen2=\dp0
  \setbox0=\hbox\bgroup
    \box0
    \ifnum\dimen0>0 \vrule width1sp height\dimen1 depth\dimen2 
    \else \vrule width1sp height1sp depth0sp\relax
    \fi\egroup
  \ht0=0pt \dp0=0pt \box0 \egroup}
\mpxshipout% line 125 ./up-Dxb.mp
-500000%
\stopmpxshipout
\mpxshipout% line 128 ./up-Dxb.mp
-400000%
\stopmpxshipout
\mpxshipout% line 131 ./up-Dxb.mp
-300000%
\stopmpxshipout
\mpxshipout% line 134 ./up-Dxb.mp
-200000%
\stopmpxshipout
\mpxshipout% line 137 ./up-Dxb.mp
-100000%
\stopmpxshipout
\mpxshipout% line 140 ./up-Dxb.mp
0%
\stopmpxshipout
\mpxshipout% line 143 ./up-Dxb.mp
100000%
\stopmpxshipout
\mpxshipout% line 146 ./up-Dxb.mp
200000%
\stopmpxshipout
\mpxshipout% line 149 ./up-Dxb.mp
300000%
\stopmpxshipout
\mpxshipout% line 152 ./up-Dxb.mp
400000%
\stopmpxshipout
\mpxshipout% line 155 ./up-Dxb.mp
500000%
\stopmpxshipout
\mpxshipout% line 159 ./up-Dxb.mp
0%
\stopmpxshipout
\mpxshipout% line 163 ./up-Dxb.mp
0.2%
\stopmpxshipout
\mpxshipout% line 167 ./up-Dxb.mp
0.4%
\stopmpxshipout
\mpxshipout% line 171 ./up-Dxb.mp
0.6%
\stopmpxshipout
\mpxshipout% line 175 ./up-Dxb.mp
0.8%
\stopmpxshipout
\mpxshipout% line 179 ./up-Dxb.mp
1%
\stopmpxshipout
\mpxshipout% line 183 ./up-Dxb.mp
1.2%
\stopmpxshipout
\mpxshipout% line 187 ./up-Dxb.mp
C$_{16}$H$_{8}$-up%
\stopmpxshipout
\mpxshipout% line 189 ./up-Dxb.mp
\Large $\mu^{abcd}(\,J\,s^{-1}\,V^{-2})$%
\stopmpxshipout
\mpxshipout% line 191 ./up-Dxb.mp
{\Large Photon Energy (eV)}%
\stopmpxshipout
\mpxshipout% line 194 ./up-Dxb.mp
D^{xx}%
\stopmpxshipout
% line 262 ./up-Dxb.mp
 \end{document}

\end{document}