### 2017-02-23 checking_maths/checking_maths.tex
Running layer responses.

\documentclass[landscape]{article}

\usepackage[spanish]{babel}
\decimalpoint
\usepackage{multirow}
\usepackage{textcomp}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{eso-pic}
\usepackage{amsmath}
\usepackage{multicol}
% \usepackage{eurosym}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.12}

\author{Reinaldo Arturo Zapata Pe\~na}
\title{Checking Maths for $\mathcal{V}^{\mathrm{ab}}$} 
\date{}

\begin{document}
\maketitle

The equation for $\mathcal{V}^{\mathrm{ab}}$ for normal incidence in the $xy$ plane
is given by
\begin{align*}
\mathcal{V}^{\mathrm{ab}} 
&= \frac{\mu^{\mathrm{abxx}}E^{2}\cos^{2}\theta + 
\mu^{\mathrm{abyy}}E^{2}\sin^{2}\theta + 
2\mu^{\mathrm{abxy}}E^{2}\cos\theta\sin\theta }
{\xi^{\mathrm{xx}}E^{2}\cos^{2}\theta + \xi^{\mathrm{yy}}E^{2}\sin^{2}\theta}, \\
&= \frac{\mu^{\mathrm{abxx}}\cos^{2}\theta + 
\mu^{\mathrm{abyy}}\sin^{2}\theta + 2\mu^{\mathrm{abxy}}\cos\theta\sin\theta }
{\xi^{\mathrm{xx}}\cos^{2}\theta + \xi^{\mathrm{yy}}\sin^{2}\theta}.
\end{align*}
For an angle $\theta = \frac{\pi}{4}$ this expression can be reduced to 
\begin{align*}
\mathcal{V}^{\mathrm{ab}} 
&= \frac{\mu^{\mathrm{abxx}} + \mu^{\mathrm{abyy}} + 2\mu^{\mathrm{abxy}}}
{\xi^{\mathrm{xx}} + \xi^{\mathrm{yy}}}.
\end{align*}

The input files are

{\tt
\begin{tabular}{lr}
chi1.kk\_xx\_yy\_zz\_\ldots &  {\rm 7 columns} \\
\hline
mu.kk\_xxxx\_xxxy\_xxxz\_xxyx\_xxyy\_xxyz\_xxzx\_xxzy\_xxzz\_\ldots & {\rm 10 columns}\\
mu.kk\_xyxx\_xyxy\_xyxz\_xyyx\_xyyy\_xyyz\_xyzx\_xyzy\_xyzz\_\ldots & \\
mu.kk\_xzxx\_xzxy\_xzxz\_xzyx\_xzyy\_xzyz\_xzzx\_xzzy\_xzzz\_\ldots & \\
mu.kk\_yxxx\_yxxy\_yxxz\_yxyx\_yxyy\_yxyz\_yxzx\_yxzy\_yxzz\_\ldots & \\
mu.kk\_yyxx\_yyxy\_yyxz\_yyyx\_yyyy\_yyyz\_yyzx\_yyzy\_yyzz\_\ldots & \\
mu.kk\_yzxx\_yzxy\_yzxz\_yzyx\_yzyy\_yzyz\_yzzx\_yzzy\_yzzz\_\ldots & \\
\end{tabular}
}

\vspace{0.5cm}
Then I pasted first each \textbf{\tt mu.kk\ldots} followed for the \textbf{\tt
chi1.kk\ldots} file to have 17 column file

\vspace{0.5cm}
{\footnotesize
\begin{tabular}{lcccccccccccccccccc}
& \multicolumn{10}{c}{\tt mu.kk\ldots} & \multicolumn{7}{c}{\tt chi1.kk\ldots} \\
\cline {2-11} \cline{13-19}
&energy & abxx & abxy & abxz & abyx & abyy & abyz & abzx & abzy & abzz & &
energy & Re xx & Im xx & Re yy & Im yy & Re zz & Im zz \\
column: & 1& 2& 3& 4& 5& 6& 7& 8& 9& 10 & & 11& 12& 13& 14& 15& 16& 17\\
\end{tabular}
}

\newpage
I defined the variables
\begin{align*}
&\texttt{coscuad} = \cos^{2}\theta \\
&\texttt{sincuad} = \sin^{2} \theta\\
&\texttt{cossin}  = \cos\theta \sin\theta 
\end{align*}
and used {\tt awk} comand to produce the corresponding \tt{\textbf{v.kk\_ab...}}
file as follows:

\noindent
{\tt awk \textquotesingle\{ print \$1, ((\$2*\textquotesingle
\,``\$coscuad''\textquotesingle + \$6* \textquotesingle\,``\$sincuad''
\textquotesingle + 2* \$3* \textquotesingle\,``\$cossin''\textquotesingle)/(\$13*
\textquotesingle\,``\$coscuad'' \textquotesingle + \$15* \textquotesingle
\,``\$sincuad'' \textquotesingle + 0.00001)) \}\textquotesingle }

\end{document}