### 2017-08-15 mu/article/article-review6-A.tex
Starting Anatoli's review.

%\documentclass[prb,11pt,tightenlines,twocolumn,aps]{revtex4-1}
\documentclass[floatfix,prb,aps,superscriptaddress,showpacs,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsmath}    % need for subequations 
\usepackage{graphicx}   % need for figures 
\usepackage{verbatim}   % useful for program listings 
\usepackage{color}      % use if color is used in text 
\usepackage{subfigure}  % use for side-by-side figures 
\usepackage{hyperref}   % use for hypertext links, including those to external 
                        % documents and URLs 
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
%\usepackage[inline]{showlabels}
% \input{/Users/bms/util/definitions}
%\input{/Users/reinaldo/Documents/tex_files/bmsdefinitions.tex}
\def\tama{10cm}
\raggedbottom           % don't add extra vertical space

\begin{document}

\title{Pure Spin Current Injection in Hydrogenated Graphene Structures}
\author{Reinaldo Zapata-Pe\~na}
\affiliation{Centro de Investigaciones en \'Optica, Le\'on,
Guanajuato 37150, M\'exico}
\author{Bernardo S. Mendoza}
\affiliation{Centro de Investigaciones en \'Optica, Le\'on,
Guanajuato 37150, M\'exico}
\author{Anatoli I. Shkrebtii}
\affiliation{University of Ontario, Institute of Technology,
Oshawa, ON, L1H 7L7, Canada}

\date{\today}

\begin{abstract}
We present a theoretical study of spin-velocity injection (SVI) of a pure spin current (PSC)
induced by  a linearly polarized light that impinges normally on the
surface of  two 50\% hydrogenated noncentrosymmetric two-dimensional 2D graphene structures. The first structure, hydrogenated at only one side, 
labeled $Up$ (also known as graphone), and the second, one labelled $Alt$, is 25\% hydrogenated at both sides. The hydrogenation opens an energy gap in both structures. We
analyze two possibilities: the first one  is by fixing the spin along a chosen
direction, and the resulting spin-velocity is calculated. Secondly,  we
choose the spin-velocity direction within (OR along?) the surface plane, and calculate its (which one) the spin-velocity injection speed
and spin orientation. This is done by changing the energy $\hbar\omega$ and
polarization angle $\alpha$  of the incoming  light. The results are calculated
within a full electronic band structure scheme using the Density Functional Theory (DFT) in the Local Density Approximation (LDA). The maxima of the spin-velocity speed are reached when
$\hbar\omega=0.084$\,eV and $\alpha=35^\circ$ for the Up structure, and $\hbar\omega=0.720$\,eV and
$\alpha=150^\circ$ for the Alt geometry. We find a speed of 668\,Km/s and 645\,Km/s for Up and
Alt structures, respectively, when the spin points perpendicularly to the surface. Also,
the response is maximized by fixing the spin-velocity direction along a high
symmetry Cartesian axis, reaching a speed of 688\,Km/s with the spin  pointing
at 13$^\circ$ from the surface normal, the for Up, and 906\,Km/s and the spin
pointing at 60$^\circ$ from the surface normal, for the Alt. These speed values are significantly 
larger than those of bulk semiconductors, such as CdSe and GaAs, thus making the
hydrogenated graphene structures excellent candidates for spintronics
applications.
\end{abstract}

\maketitle

\section{Introduction}
\label{sec:introduction}

Spintronics is an emerging research field of electronics in which the
manipulation and transport of spin of electrons in a solid state materials  is crucial (central), adding a new degree of freedom to conventional charge
manipulation.\cite{wolfSC04,fabianAPS07} 
At present, there is an increasing interest in attaining the same level of
control over the transport of spin at micro- or nano-scales, as it has been done
for the flow of charge in typical 3D-bulk based electronic devices.\cite{awschalomNP2007} Several
semiconductor spintronics devices have been proposed \cite{majumdarAPL06,
dattaAPL90,gotteNat16,pershinPRB08}, and some of them require spin polarized
electrical current \cite{awschalomSSBM13} or pure spin current (PSC).
One of the difficulties  in achieving (creating) measurable  the of spin current and development of PSC based
semiconductor devices is the fact that the spin relaxation time in conventional 
semiconducting materials is short to enable the spin transport, thus resulting
in a non-observable spin current.\cite{murakamiSc03}
For PSC there is no net motion of charge; spin-up electrons move in a given
direction, while spin-down electrons travel in the opposite one. This effect can be due to one-photon absorption of linearly polarized light by a
semiconductor, with filled valence bands and empty conduction bands, illuminated
by light with photon energy larger than the energy gap. This phenomenon can
be due to spin injection,\cite{malPRB03} Hall Effects,\cite{sinovaPRB04}
interference of two optical beams,\cite{bhatPRL00, najmaiePRB03} or one photon
absorption of linearly polarized light\cite{bhatPRL05}. The last effect has been observed in
gallium arsenide (GaAs),\cite{zhaoPRL2006, stevensPRL03} aluminum-gallium
arsenide (AlGaAs),\cite{stevensPRL03} and Co$_2$FeSi.\cite{kimuraNGPAM12}

Graphene, an allotrope of carbon with hexagonal 2D lattice structure, demonstrates 
properties such as fractional quantum Hall effect at room temperature, excellent
thermal transport properties, excellent conductivity\cite{heerscheNat07} and
strength \cite{geimNM07, reinaNL08, novoselov2S07, balandinNL08}, being a
perfect platform  for two-dimension (2D) electronic systems; however, most
electronic applications are disabled by the absence of a semiconducting gap.
Recent studies demonstrate that a narrow band gap  can be opened in graphene by
applying an electric field,\cite{zhangN09} reducing the surface
area,\cite{hanPRL07} or applying uniaxial strain.\cite{niACSN08} Another
possibility to open the gap is by doping; this has been successfully achieved
using nitrogen,\cite{weiNL2009} boron-nitrogen,\cite{guoIJ11}
silicon,\cite{colettiPRB10} noble-metals,\cite{varykhalovPRB10} and
hydrogen.\cite{eliasS09, guisingerNL09, samarakoonACSN10}
Depending on the percentage of hydrogenation and spatial arrangements of
the hydrogen-carbon bonds, hydrogenated graphene demonstrates  different structural  
configurations and a tunable electron gap, as it has been proven in [1].
In this paper, we present two 50\% hydrogenated graphene noncentrosymmetric
structures, both presenting a discernible band gap. The first one, labelled as the Up structure (also known as graphone, see [2] and Refs therein), shown in
Fig. \ref{fig:up-struc}. Graphone has hydrogen atoms bonded to the carbon layer only on
the upper side of the structure, we consider here the magnetic isomer of graphone, with the so-called “chair” structure. In contrast,  the Alt structure, shown in Fig. 
\ref{fig:alt-struc}, has hydrogen alternating on the upper and bottom sides of 
the carbon sheet.\cite{zapataPSB2016}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/upstruc2}
    \\
    \includegraphics[width=\tama]{figures/upstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Up
      structure along with the 
      Cartesian $xyz$ directions. The dark (light) spheres are the C (H) atoms. The primitive hexagonal unit cell is also shown.
}
    \label{fig:up-struc}
\end{figure}
\begin{figure}[ht!]
    \centering
    \includegraphics[width=\tama]{figures/altstruc2}
    \\
    \includegraphics[width=\tama]{figures/altstruc1}
    \caption{Side (top panel) and top (bottom panel) views of the Alt
      structure along with the 
      $xyz$ directions. The dark (light) spheres are the C (H) atoms. The primitive rectangular unit cell is also displayed.

}
    \label{fig:alt-struc}
\end{figure}

Using those structures, we address theoretically  the spin velocity
injection (SVI) by one-photon absorption of linearly polarized light. Since wehave 2D structures, we do the analysis for two cases (not clear to me this sentence). The first case is by fixing the
spin of the electrons along vertical to the surface direction (?)$z$, , with the velocity
directed onto (is this parallel or not?) the surface of the structure in the $xy$ plane. In the second 
case we fix the spin velocity in the $x$ or $y$ direction and the resulting spin is then directed outward of the $xy$ plane. % $xyz$. (Do we need to repeat “Cartesian” many times?)
The SVI is an optical effect that quantifies the velocity, at which a PSC moves
along the direction $\bf{a}$ with the $\bf{a}$ components to be determined,  % ####???, $\mathrm{a}$ and 
with the spin of electron polarized
along the direction $\bf{b}$. % $\mathrm{b}$. One-photon absorption of linearly
polarized light produce an even distribution of electrons in $\mathbf{k}$
space, regardless of the symmetry of the material, resulting in a null
electrical current.\cite{bhatPRL05} Then, the electrons excited to the
conduction bands at opposite $\mathbf{k}$ points will result in opposite spin
polarizations producing no net spin injection in centrosymmetric materials (?).\cite{bhatPRL05} If the
crystalline structure of the material is noncentrosymmetric, the spin
polarization injected at a given $\mathbf{k}$ point not necessarily
vanishes,\cite{alvaradoPRL85, schmiedeskampPRL88}. Therefore, since the velocities
of electrons at opposite $\mathbf{k}$ points are opposite, a PSC will be
produced. Since both the structures presented here are noncentrosymmetric, they are
good candidates in which this effect can be induced.

This paper is organized as follows. In Section \ref{sec:theory} we present the
formalism and the main expressions  that describe PSC and SVI. In Section \ref{sec:results} we
describe the numerical details  and  discuss the corresponding SVI spectra for the
Up and Alt structures. Finally, we summarize our finding in
Section \ref{sec:conclusions}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%               %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  T H E O R Y  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%               %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Theory} % (fold)
\label{sec:theory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% Theory: Spin velocity %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this section, we  summarize the theoretical approach, involved in 
calculation of the spin velocity
injection (SVI) resulting from the
pure spin current (PSC).

%%%BMSD
The operator that describes the electronic SVI is written as
\begin{align}\label{z.1}
\hat K^{\mathrm{a}\mathrm{b}}=\frac{1}{2}\left( \hat v^\mathrm{a} \hat S^\mathrm{b} 
+\hat  S^\mathrm{b} \hat v^\mathrm{a}\right) 
,
\end{align} 
where $\hat{\mathbf v}=[\hat{\mathbf r},\hat H_0]/i\hbar$ 
is the velocity operator, with
$\hat {\mathbf r}$ 
%\end{document}
being the position operator and $\hat H_0$ the unperturbed
ground state Hamiltonian;
the Roman superscripts  indicate Cartesian coordinates. 
To obtain the expectation value of 
$\hat K^{\mathrm{a}\mathrm{b}}$, we use the length gauge for the perturbing
Hamiltonian, written as
\begin{align}\label{z.2}
\hat H_{\text{p}}=-e\hat{\mathbf r}\cdot{\mathbf E}(t)
,
\end{align}   
where the applied electric field of the laser is given by
\begin{align}\label{z.3}
{\mathbf E}(t)= {\mathbf E}(\omega)e^{-i\omega t} + {\mathbf E}^*(\omega)e^{i\omega t}
.
\end{align}
In order to 
calculate the response of the system to ${\mathbf E}(t)$, one needs to
take into account the excited coherent superposition
of the spin-split conduction bands inherent to the 
noncentrosymmetric 
semiconductors considered in this work.
%Although, this splitting is small [15,16], is required to obtain the
%correct results\cite{nastosPRB07}.
% The conduction bands in the 
% noncentrosymmetric  
% semiconductors  
% are spin split by a small amount,[15,16] typically
%   smaller than the energy width of the laser pulse, and so the pulse
%   excites a coherent superposition of the two conduction bands. Even
%   for very long pulses with narrow energy widths, dephasing effects
%   lead to an energy width of the bands large enough that spin-split
%   states can become quasi-degenerate. 
To include the coherence, we follow Ref.~\onlinecite{nastosPRB05} and
use a multiple
scale approach that solves the equation of motion for the single
particle density matrix ${\rho}_{mn}({\mathbf k};t)$, leading to
\begin{align}\label{z.4}
&\frac{\partial \rho_{cc'}({\mathbf k})}{\partial t} =
\frac{e^{2}E^{\mathrm{a}}(\omega)E^{\mathrm{b*}}(\omega)}
{i \hbar^{2}}
\sum_{v}r^{\mathrm{a}}_{cv}({\mathbf k}) r^{\mathrm{b}}_{vc'}({\mathbf k})
\left( \frac{1}{\omega - \omega_{c'v}({\mathbf k}) - i \epsilon} 
- 
\frac{1}{\omega - \omega_{cv}({\mathbf k}) + i \epsilon} \right)
,
\end{align}
where we assumed that the conduction bands $c$ and $c'$
are quasi-degenerate  states, and we take $\epsilon\to 0$ at the end of the
calculation.  The spin-splitting of the valence ($v$) bands is very
small, and is neglected throughout this work.\cite{nastosPRB07}
The matrix elements of any operator ${\cal O}$ are given by
${\cal O}_{nm}({\mathbf k})=\langle n{\mathbf k}|\hat{\cal O}| m{\mathbf k}\rangle$,
where 
$H_{0}|n{\mathbf k}\rangle = \hbar \omega_{n}({\mathbf k})|n{\mathbf k}\rangle$ 
with $\hbar \omega_{n}({\mathbf k})$ being the energy of the
electronic band $n$ at point ${\mathbf k}$ in the irreducible Brillouin zone
(IBZ),  $|n{\mathbf k}\rangle$ is the Bloch state, and 
$\omega_{nm}({\mathbf k})=\omega_{n}({\mathbf k})-\omega_{m}({\mathbf k})$.
Using 
$\mathcal{O} = \mathrm{Tr}(\hat{\rho}\hat{\mathcal{O}})$
for the expectation value of an observable $\mathcal{O}$, 
where $\mathrm{Tr}$ denotes the trace, we obtain
\begin{align}\label{z.5}
\mathcal{O} 
%=&
%\int \frac{d^{3}k}{8\pi^{3}}\sum_{c} \langle c {\mathbf k} 
%| \hat{\rho} \hat{\mathcal{O}} |
%c{\mathbf k} \rangle \nonumber \\
=& 
\int \frac{d^{3}k}{8\pi^{3}} \sum_{cc'} \rho_{cc'}({\mathbf k}) 
\mathcal{O}_{c'c}({\mathbf k}),
\end{align}
where we used 
the closure
relationship $\sum_{n}|n{\mathbf k}\rangle \langle n{\mathbf k}| = 1$,
where $n=v,c$,
and the fact that $\rho_{vn}({\mathbf k})=\rho_{nv}({\mathbf k})=0$ for $n=c,c'$. 
Therefore, 
using  Eqs. \eqref{z.4} and \eqref{z.5},
the
rate of change of $\mathcal{O}$,
$\dot{\mathcal{O}} 
=
\mathrm{Tr} \left( \dot{\hat\rho} \hat{\cal O} \right)
$,
 is given by
\begin{align}
\dot{\mathcal{O}} 
&=\frac{e^{2}}{i\hbar^{2}} \int \frac{d^{3}k}{8\pi^{3}} 
\sum'_{cc'} \mathcal{O}_{c'c}({\mathbf k}) 
r^{\mathrm{a}}_{cv}({\mathbf k})  r^{\mathrm{b}}_{vc'}({\mathbf k})  
\left( \frac{1}{\omega - \omega_{c'v}({\mathbf k})  - i\epsilon} - 
\frac{1}{\omega - \omega_{cv}({\mathbf k})  + i\epsilon} \right)
E^{\mathrm{a}}(\omega) E^{\mathrm{b*}}(\omega)
\label{eq:dotO}
.
\end{align}
The prime symbol $'$ in the sum means that
$c$ and $c'$ are quasi-degenerate states, and the sum only covers these states.

Replacing  $\hat{\cal O} \rightarrow \hat{K}^{\mathrm{ab}}$, in 
the above expression, one can show that
\begin{equation}
\dot{K}^{\mathrm{ab}}(\omega) =
\mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega),
\label{eq:dotk}
\end{equation}
where repeated Cartesians are summed, and 
\begin{equation}\label{eq:mu}
\begin{aligned}
\mu^{\mathrm{abcd}}  (\omega) &
=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}} \sum'_{vcc'}
  \delta(\omega-\omega_{cv}({\mathbf k}) 
\mathrm{Re} & \left[ K^{\mathrm{ab}}_{cc'}({\mathbf k}) 
\left(  
r^{\mathrm{c}}_{vc'}({\mathbf k})   
r^{\mathrm{d}}_{cv }({\mathbf k})  +
(c \leftrightarrow d)  
\right) 
\right]
\end{aligned}
\end{equation} 
is the pseudotensor that describes the rate of change of the  PSC process in
semiconductors. To derive what we presented above we used
$
K^{\mathrm{ab}}_{nm}({\mathbf -k}) = K^{\mathrm{ab*}}_{nm}({\mathbf k}) 
$
which follows from time-reversal invariance. 
Since $\mu^{\mathrm{abcd}}(\omega)$ is real, we have that
$\mu^{\mathrm{abcd}}(\omega) =
\mu^{\mathrm{abdc}}(\omega)$. 
We point out that Eq.~\eqref{eq:mu} 
is –identical to Eq. (3) of Bhat et al.\cite{bhatPRL05},
derived using the semiconductor optical Bloch equations.
Using the closure relation,
\begin{equation}
K^{\mathrm{ab}}_{cc'}({\mathbf k}) = \frac{1}{2}
\sum_{l=v,c}
\left(v^{\mathrm{a}}_{cl}({\mathbf k})S^{\mathrm{b}}_{lc'}({\mathbf k})
+S^{\mathrm{b}}_{cl}({\mathbf k}) v^{\mathrm{a}}_{lc'}({\mathbf k})
\right)
,
\label{eq:velspimatelem}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsection{Spin velocity injection} % (fold)
%\label{sec:theory-pure_spin_current}
we define the spin velocity injection (SVI) as
\begin{equation}\label{eq:vab-w}
\mathcal{V}^{\mathrm{ab}}(\omega) \equiv
\frac{\dot{K}^{\mathrm{ab}}(\omega)}{(\hbar/2) \dot{n}(\omega)},
\end{equation}  
which gives the velocity, along direction $a$, % $\mathrm{a}$, 
at which the spin
moves in a polarized state along direction $b$. % $\mathrm{b}$. 
The carrier injection rate
$\dot n(\omega)$ is written as\cite{nastosPRB05}
\begin{equation}
\dot{n}(\omega) =
\xi^{\mathrm{ab}}(\omega) E^{c }(\omega) E^{d*}(\omega)
\label{eq:dotn}
\end{equation}
where the tensor 
\begin{equation}\label{eq:xi}
\begin{aligned}
\xi^{\mathrm{ab}}(\omega)
&
=
\frac{2\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}k}{8 \pi^{3}}
 \sum_{vc}
r^{\mathrm{a}}_{vc'}({\mathbf k})  
r^{\mathrm{b}}_{cv }({\mathbf k})  
\delta(\omega-\omega_{cv}({\mathbf k})), 
\end{aligned}
\end{equation}
is related to the imaginary part of the linear optical 
response tensor by
$\mathrm{Im}[\epsilon^{\mathrm{a}\mathrm{b}}(\omega)]=2\pi\epsilon_0\hbar\xi^{\mathrm{a}\mathrm{b}}(\omega)$.

The function ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$ allows us to quantify two very
important aspects of PSC. On one hand, we can fix the spin direction
along $\mathrm{b}$  
and calculate the resulting electron velocity.  
On the other hand, we can fix the velocity of the electron
along $\mathrm{b}$ and study the resulting direction along which the
spin is polarized.
To this end, the additional advantage of  2D structures, besides being
 noncentrosymmetric, is that we can use an incoming linearly
polarized  light
at normal incidence, and use the  direction of the polarized  electric
field to control ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$.
Indeed, writing 
${\mathbf E}(\omega) = E_0(\omega)(\cos\alpha\,\hat{\mathbf x}+\sin\alpha\,\hat{\mathbf y})$,
where $\alpha$ is the polarization angle, we obtain from
 Eq. \eqref{eq:vab-w}
that
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= 
\frac{2}{\hbar\xi(\omega)}
\left(\mu^{\mathrm{abxx}}(\omega)\cos^{2}\alpha + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}\alpha + 
\mu^{\mathrm{abxy}}(\omega)\sin 2\alpha\right)
,
\label{eq:vab-aw}
\end{align}
\end{widetext}
for the structures chosen in this article,
$\xi^{\mathrm{xx}}(\omega)=\xi^{\mathrm{yy}}(\omega)\equiv\xi(\omega)$, and $\xi^{\mathrm{xy}}(\omega)=0$.
Next, we identify two options for ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$.

\subsection{Fixing the spin polarization}\label{sec:theory-fixspin}

Analyzing the SVI, Eq. \eqref{eq:vab-aw}, we define (calculate?) the magnitude of
the electron velocity along the plane of the structure,
with the spin polarized along $\mathrm{b}$ direction as
\begin{equation}
\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)
\equiv
\sqrt{
[\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\end{equation}
and define the angle at which the velocity is directed on the $xy$ plane as
\begin{equation}
\gamma_{\sigma^\mathrm{b}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{yb}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{xb}}(\omega,\alpha)} \right)
.
\label{eq:gamma-ang}
\end{equation}
% where this angle is measured in the counter-clockwise direction from the
% positive $x$. 
We also define two special angles
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\parallel(\omega,\alpha) = \alpha, 
\label{eq:gamma-par} 
\end{equation}
and
\begin{equation}
\gamma_{\sigma^\mathrm{b}}^\perp(\omega,\alpha) = \alpha \pm 90^{\circ},
\label{eq:gamma-perp}
\end{equation}
corresponding to the electron velocity being parallel or perpendicular
to the incoming light polarization, 
respectively. The subscript $\sigma^\mathrm{b}$ denotes the spin along $\mathrm{b}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Theory: Fixing vel %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Fixing the electron velocity.}\label{sec:theory-fixvel}

Fixing the calculated velocity along $\mathrm{a}=x$ or $\mathrm{a}=y$,
we define its corresponding magnitude as
\begin{align}
\mathcal{V}_{\mathrm{a}}(\omega,\alpha) \equiv 
\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
from where we see that the spin would be oriented
in the $xyz$ system's coordinates
defined with a polar angle
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
= 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{\mathcal{V}_{\mathrm{a}}(\omega,\alpha)} \right),
 \quad 0 \leq \theta \leq \pi, 
\label{eq:polar-ang}
\end{align}
and an azimuthal angle
\begin{align}
\varphi_{\mathrm{a}} (\omega,\alpha)
=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\quad 0 \leq \varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%   R E S U L T S   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%                   %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Results} % (fold)
\label{sec:results}

\begin{table}[t]
\center
\begin{tabular}{ccccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H & -0.61516 & -1.77416 &  0.73196 \\
H &  0.61518 &  0.35514 &  0.73175 \\
C & -0.61516 & -1.77264 & -0.49138 \\
C & -0.61516 & -0.35600 & -0.72316 \\
C &  0.61516 &  0.35763 & -0.49087 \\
\hline
\end{tabular}
\caption{Atomic positions in the unit cell of Up structure as the structural input for our calculations (the primitive hexagonal unit cell contains two C and one H atoms, see Fig. 1). ADD EXTRA C atom
}
\label{tab:up-unitcell}
\end{table}
% 
% 
\begin{table}[t]
\center
\begin{tabular}{cccc}\\
\hline
\quad Atom \qquad & \multicolumn{3}{c}{Position (\AA)} \\
\cline{2-4}
\quad type \qquad & $x$ & $y$ & $z$  \\
\hline
H &  -0.61516 &  -1.42140 & \ 1.47237 \\
C &  -0.61516 &  -1.73300 & \ 0.39631 \\
C & \ 0.61516 & \ 1.73300 & \ 0.15807 \\
C & \ 0.61516 & \ 0.42201 &  -0.15814 \\
C &  -0.61516 &  -0.37396 &  -0.39632 \\
H &  -0.61516 &  -0.68566 &  -1.47237 \\
\hline
\end{tabular}
\caption{Atomic positions in the unit cell of Alt structure, which coincides with its primitive cell. 
}
\label{tab:alt-unitcell}
\end{table}

We present the calculated results of $\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)$ and
$\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ for the C$_{16}$H$_{8}$-Up  and
C$_{16}$H$_{8}$-Alt structures, both noncentrosymmetric 2D carbon
systems with 50\% hydrogenation, which are differently structurally arranged. We remind
that the Up structure has hydrogen atoms only on the upper side of the carbon
sheet, while the Alt structure has alternating hydrogen atoms on the upper and
bottom sides. We take the carbon lattice to be along the $xy$ plane
for both structures, and the carbon-hydrogen bonds are outward the plane, they are perpendicular to  $xz$
plane for the Up structure, as depicted in Figs. \ref{fig:up-struc}, and off the normal for the Alt structure \ref{fig:alt-struc}. The
coordinates for the Up and Alt unit cells of the structures are presented in
Tables \ref{tab:up-unitcell} and \ref{tab:alt-unitcell}.

We calculated the self-consistent ground state and the Kohn-Sham states within 
density functional theory in the local density approximation (DFT-LDA), with a
planewave basis using the ABINIT code \cite{gonzeCPC09}.
% 
We used Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space
Gaussian pseudopotentials \cite{hartwigsenPRB98}, including the spin-orbit
interaction needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented
in Eq. \eqref{eq:mu}.
% 
The convergence parameters for the calculations,  corresponding to
the Up and Alt structures are cutoff energies of 40\,Ha and 65\,Ha, resulting
in LDA energy band gaps of 0.084\,eV and 0.718\,eV, respectively. The energy
eigenvalues and matrix elements for the Up and Alt structures were calculated
using 12802 ${\mathbf k}$ points and 14452 ${\mathbf k}$ points in the IBZ; to
integrate $\mu^{\mathrm{abcd}}(\omega)$ and $\xi^{\mathrm{a}\mathrm{b}}(\omega)$ the
linearized analytic tetrahedron method (LATM) has been used.\cite{nastosPRB07} We neglect the
anomalous velocity term $\hbar(\boldsymbol{\sigma}\times\nabla V)/4m^2c^2$, where $V$ is the
crystal potential, in $\hat{\mathbf v}$ of Eq.~\eqref{z.1}, as this term is known to
give small contribution to PSC.\cite{bhatPRL05} Therefore,
$[\hat{\mathbf v},\hat{\mathbf S}]=0$, and Eq.~\eqref{z.1} reduces to $\hat K^{\mathrm{a}\mathrm{b}}=\hat
v^\mathrm{a} \hat S^\mathrm{b}=\hat S^\mathrm{b} \hat v^\mathrm{a}$. Finally, the prime in the sum of
Eq.~\eqref{eq:mu} is restricted to quasi-degenerated conduction bands $c$ and
$c'$ that are closer than 30 meV, which  is both typical laser-
pulse energy width and the thermal room-temperature energy level broadening.\cite{nastosPRB07}
% 
% It is known that the DFT with LDA approximation misestimates the energy band
% gap of semiconductors. To correct this one has to include the many-body
% interaction using i.e. the so-called GW approximation but this technique has a
% very high computational cost and is out of the scope in this paper.
% Nevertheless, DFT still remains a mainstream tool for calculating diverse
% properties derived from the electronic band structure.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% Results: Spin velocity %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{SVI: Spin velocity injection} % (fold)
\label{sec:res-spin_velocity}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%% Res: mu & V^{ab} comparison %%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\centering
\includegraphics[width=\tama]{plots/2-vab-str-comp}
\caption{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ 
  vs. $\hbar\omega$, for the angles $\alpha$ that maximize the signal. The largest velocity are at the low
  energy regions of the spectra for the Alt and Up structures, starting (going up?) at the energy gap of each
  structure. In the high energy regions, the values of
  $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ are also very large compared to the 3D case of CdSe and GaAsThe, shown at the bottom panel . } 
\label{fig:vab-str-comp}
\end{figure}

In Fig. \ref{fig:vab-str-comp}, we show $\mathcal{V}^{\mathrm{ab}}
(\omega,\alpha)$ vs. $\hbar\omega$ for the directions $\mathrm{a}\mathrm{b}$ (confusing to me), and for the
angle $\alpha$, at which the signal is 
maximized , for the Up  and Alt structures, and for CdSe and
GaAs bulk systems, shown for comparison.
% 
As expected from Eq.~\eqref{eq:mu},
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ rises from zero right at the corresponding energy gap of each
system. For the 2D structures considered, the spectrum contains two narrow energy regions
with strong response, while for the bulk systems, the spectra covers a
rather wide energy range, but with a much weaker response.
% 
For the Up structure, at $\mathrm{a}\mathrm{b}=yz$ and $\alpha=35^\circ$ the response is maximized,
which means that an incoming light with its electric field polarized at
35$^\circ$ from the $x$ direction will induce electrons to move along $y$
(parallel to the surface), with their spin polarized along $z$ (perpendicular to
the surface). The corresponding speeds are the following:
% 
Right at the energy onset, ${\cal V}^{yz}(\omega,\alpha)=668$\,Km/s, remains
almost constant for 65 meV, and then decreases to zero. A second region with high velocity is 
above 1.946 eV with two, opposite in sign, maxima of the speed:
${\cal V}^{yz}(\omega,\alpha)=266.3$ Km/s at $\hbar\omega=1.954$\,eV, and
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)=-241.4$ Km/s at $\hbar\omega=1.958$\,eV; a positive
(negative) ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ means that the electrons move parallel
(antiparallel) to the electric field.
Likewise, for the Alt structure, we also find that $\mathrm{a}\mathrm{b}=yz$ and
$\alpha=150^\circ$ maximizes the response, where two extreme values of
${\cal V}^{\mathrm{y}\mathrm{z}}$ are found, one at  $\hbar\omega=0.720$\,eV of
${\cal V}^{\mathrm{y}\mathrm{z}}=-711.9$ Km/s, and the other at $\hbar\omega=0.911$\,eV of
${\cal V}^{\mathrm{y}\mathrm{z}}=-330.6$ Km/s.

For the bulk structures, we calculate ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega)$ from
Eq.~\eqref{eq:vab-w} by simply using $\boldsymbol{\mu}_{\mathrm{max}}$. For CdSe, we find
that for $\hbar\omega=0.844$\,eV, $\boldsymbol{\mu}_{\mathrm{max}}\to \mu^{zzzz}$, and
${\cal V}^{\mathrm{z}\mathrm{z}}(\omega)=-59.0$ Km/s, and for GaAs at $\hbar\omega=2.324$\,eV,
$\boldsymbol{\mu}_{\mathrm{max}}\to\mu^{\mathrm{aaaa}}$ and ${\cal V}^{\mathrm{a}\mathrm{a}}(\omega,\alpha)=-28.7$\,Km/s, with
$\mathrm{a}=x,y,z$. For these bulk semiconductors, the $x$, $y$, and $z$ axis are
taken along the standard cubic unit cell directions, [100], [010], and [001],
respectively. In Table \ref{tab:vab-str-comp}, we compare 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega,\alpha)$ for the 2D structures considered and bulk crystals.
% 
We stress that, as shown in the figure, for 
the 2D structures have maxima in 
${\cal V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$  
are higher
than for the bulk crystals. In particular, the Alt
structure demonstrates a ${\cal V}^{\mathrm{a}\mathrm{b}}(\omega;\alpha)$ more than ten times larger than
that of CdSe and GaAs.
\begin{table}%[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
System \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& type & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
$Up$    & 2D   & 35    & 0.084  & $\mathrm{yz}$ &  660.5 \\
      &      &       & 1.954  & $\mathrm{yz}$ &  266.3 \\
      &      &       & 1.958  & $\mathrm{yz}$ & -241.4 \\
$Alt$   & 2D   & 150   & 0.720  & $\mathrm{yz}$ & -711.9 \\
      &      &       & 0.911  & $\mathrm{yz}$ & -330.6 \\
CdSe  & bulk & --    & 0.844  & $\mathrm{zz}$ &  -59.0 \\
GaAs  & bulk & --    & 2.324  & $\mathrm{xx}$ &  -28.7 \\
\hline
\end{tabular}
\caption{Comparison of the reported maximum values of
$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$ for the different structures and their
corresponding polarization angle $\alpha$ and energy $\hbar\omega$ .
}
\label{tab:vab-str-comp}
\end{table}

\subsection{Fixing spin} % (fold)
\label{sec:res-fixspin}

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{upplots/up-vsz-w1}
\caption{Color online. For the Up structure, the top panel shows 
${\cal V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ SvsS. $\hbar\omega$ and $\alpha$, and the bottom panel
shows $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red line, long dashes?), and
${\cal V}_{\sigma^\mathrm{z}}(\omega,\alpha)$ (left scale, black solid line?), $vs$. $\alpha$, for
$\hbar\omega=0.084$\,eV, i.e. along the ridge shown in the 3D plot. }
\label{fig:up-vsz-w1}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{upplots/up-vsz-w2}
\caption{For the Up structure, the top panel shows ${\cal V}_{\sigma^\mathrm{z}} (\omega,\alpha)$
vs. $\hbar\omega$ and $\alpha$, and the bottom panel shows $\gamma_{\sigma^\mathrm{z}}
(\omega,\alpha)$ (right scale, red line, long dashes), and ${\cal V}_{\sigma^\mathrm{z}} (\omega,\alpha)$ (left
scale, black solid line), vs. $\alpha$, for $\hbar\omega=1.954$\,eV, i.e. along the
highest ridge shown in the 3D plot. }
\label{fig:up-vsz-w2}
\end{figure}

In this subsection, we calculate ${\cal V}_{\sigma^\mathrm{z}}(\omega,\alpha)$, Eq.
\eqref{eq:vs-mag}, for the case with the spin fixed along $z$, i.e.,
directed perpendicularly to the surface of the Up and Alt structures. Also, we
calculate $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ from Eq. \eqref{eq:gamma-ang}, which
determines the direction  of the injected electrons movement along the
surface of  each structure. We mention that we have also  analyzed 
the cases when the spin  is directed along $\mathrm{x}$ or $\mathrm{y}$,
finding similar qualitative results to those presented below.
%

\subsubsection{Up structure}\label{up:fs}

In the top panel of Fig. \ref{fig:up-vsz-w1}, we plot
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ vs.
0.080\,eV$\leq\hbar\omega\leq$0.096\,eV (similar energy range for the Up
structure shown in the left panel of Fig.~\ref{fig:vab-str-comp}) and
$0^\circ\leq\alpha\leq 180^\circ$.
% 
We see a broad peak that reaches the maximum of $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) =
739.7$\,Km/s at $\alpha=35^{\circ}$ and $\hbar\omega=
0.084$\,eV,  The variation of $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha)$ as a function of $\alpha$, which comes from the interplay of the
$\boldsymbol{\mu}$ tensor components as multiplied by the  trigonometric functions of
Eq.~\eqref{eq:vab-aw}, gives a sizable set of values between 739.7\,Km/s and
165.4\,Km/s, for 0.084\,eV$\leq\hbar\omega\leq$0.090\,eV. 
% 
In the bottom panel, we show $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha)$ $vs.$ $\alpha$ (left scale, black line), at $\hbar\omega= 0.084$\,eV,
thus following the ridge in the 3D plot of the top panel. 
Also,
we plot the
corresponding velocity angle $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$
(right scale, red line),
where it is very interesting to see that $\gamma_{\sigma^z}(\omega,\alpha)$ is centered
at 64.55$^\circ$ with a rather small deviation of only $\pm 0.03^\circ$,
for the whole range of $\alpha$. This result means that for $\hbar\omega=0.084$ eV
and for all values of $\alpha$, the electrons, with the chosen spin pointing along
$z$, will move at the angle of $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha) \sim
64.5^{\circ}$ with respect to the $x$ direction, with the range of  high
speeds $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ shown in the figure.
% 
Also, from Eq. \eqref{eq:gamma-par} we find that
$\gamma^\parallel_{\sigma^\mathrm{z}} (\omega,\alpha)=\alpha = 64.56^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 631.1$\,Km/s (see green
arrow), and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=64.50^\circ$,
gives $\alpha=154.50^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 191.5$ Km/s (see blue
arrow). Thus, at $\hbar\omega=0.084$ eV, an incident field, polarized at $\alpha
\sim 65.5^\circ$ or $\sim 154.5^\circ$, injects electrons with their spin
polarized along $z$, which move parallel or perpendicular to the incident
electric field,  with a speed of 631.14\,Km/s or 191.5\,Km/s, respectively.

Now, we analyze the results for the second energy range of the Up structure
shown in the top right panel of Fig.~\ref{fig:vab-str-comp}. In the top panel
of Fig. \ref{fig:up-vsz-w2}, we plot $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha)$ vs. 1.950\,eV$\leq\hbar\omega\leq$1.960\,eV and
$0^\circ\leq\alpha\leq 180^\circ$.
% 
We see two broad peaks that maximize at $\alpha=35^{\circ}$ and $\hbar\omega=
1.954$\,eV, with a value of $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) =
193.5$\,Km/s, and at $\alpha=35^{\circ}$ and $\hbar\omega= 1.957$\,eV, with a value
of $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 170.6$\,Km/s. 
% 
We only analyze the highest  maximum in the bottom panel, where we  show
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ vs. $\alpha$ (left scale,
black line), at $\hbar\omega= 1.954$\,eV, thus following the highest ridge shown
in the 3D plot of the top panel. Also, we plot the corresponding
velocity angle $\gamma_{\sigma^\mathrm{z}}(\omega,\alpha)$ (right scale, red
line),
where in
this case we see that the values of $\gamma_{\sigma^z}(\omega,\alpha)$ have more
dispersion, as a function of $\alpha$, than for the lower energy range shown in
the bottom panel of Fig.~\ref{fig:up-vsz-w1}.
% 
However, $\gamma_{\sigma^z}(\omega,\alpha)\sim 77.8^\circ$ is nearly constant from
$\alpha=0^\circ$ up to $\alpha\sim 85^\circ$. In this case, we find that
$\gamma^\parallel_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha=78.0^\circ$, with
$\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) = 115.0$\,Km/s (as indicated by green
arrow), and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=167.8^\circ$, gives
$\alpha=77.8^\circ$, with $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) =
65.6$\,Km/s (see blue arrow). 
% 
Thus, through the correct choice of $\hbar\omega$ and $\alpha$ we could inject
electrons, in this case with their spin polarized along $z$, which move parallel
or perpendicular to the incident electric field, with finite speeds (what does “finite” mean?).

% the maxima for
% $\hbar\omega=0.084$ eV and $\alpha=35^{\circ}$, with
% $\mathcal{V}_{\sigma^{\mathrm{x}}}(\omega,\alpha)=219.6$ Km/s and
% $\mathcal{V}_{\sigma^{\mathrm{y}}}(\omega,\alpha)=280.3$ Km/s,
% $\gamma_{\mathrm{x}}(\omega,\alpha) = 79.3^{\circ}$, and 
% $\gamma_{\mathrm{y}}(\omega,\alpha) = 78.2^{\circ}$.

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{altplots/alt-vsz}
\caption{Color online. For the Alt structure, the top panel shows ${\cal V}_{\sigma^\mathrm{z}}
(\omega,\alpha)$ vs. $\hbar\omega$ and $\alpha$, and the bottom panel shows
$\gamma_{\sigma^\mathrm{z}} (\omega,\alpha)$ (right scale, red line), and ${\cal V}_{\sigma^\mathrm{z}}
(\omega,\alpha)$ (left scale, black line), vs. $\alpha$, for $\hbar\omega=0.720$\,eV, i.e.
along the ridge shown in the 3D plot.}
\label{fig:alt-vsz}
\end{figure}

\subsubsection{Alt structure}

We proceed to analyze the Alt structure, just as we did with the Up structure, but
in this case, we only choose the lower energy range shown in the left central
panel of Fig.~\ref{fig:vab-str-comp}. 
% 
In the top panel of Fig. \ref{fig:alt-vsz}, we plot
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha)$ vs.
0.715\,eV$\leq\hbar\omega\leq$0.725\,eV and $0^\circ\leq\alpha\leq 180^\circ$. We
see a broad peak that maximizes at $\alpha=150^{\circ}$ and $\hbar\omega=
0.720$\,eV, with a value of $\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha) =
644.9$\,Km/s.
% 
In the bottom panel, we  show $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha)$ vs. $\alpha$ (left scale, black line), at $\hbar\omega= 0.720$\,eV,
 thus following the highest ridge shown in the 3D plot of the top
panel. Also, we plot the corresponding velocity angle $\gamma_{\sigma^\mathrm{z}}
(\omega,\alpha)$ (right scale, red line), where now we see that
$\gamma_{\sigma^z}(\omega,\alpha)$ is centered at $109.2^{\circ}$ having
variations of $\pm 1.0^{\circ}$ for $0^{\circ} \leq
\alpha \leq 180^{\circ}$.
% 
In this case, we find that $\gamma^\parallel_{\sigma^\mathrm{z}} (\omega,\alpha) =
\alpha = 108.8^\circ$, with $\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) =
450.05$\,Km/s (see green arrow), and that from Eq. \eqref{eq:gamma-perp},
$\gamma^\perp_{\sigma^\mathrm{z}}(\omega,\alpha)=\alpha-90^\circ=110.0^\circ$,
gives $\alpha=20.0^\circ$, with $\mathcal{V}_{\sigma^{\mathrm{z}}}
(\omega,\alpha) = 60.84$\,Km/s (see blue arrow). Thus, as for the Up structure,
we could inject electrons with a fixed spin, which moves parallel or perpendicular
to the incident electric field.

% Again, for the cases in which the spin polarization is parallel to the
% surface the maximum values are
% $\mathcal{V}_{\sigma^{\mathrm{x}}}(\omega,\alpha) = 297.8$ Km/s and
% $\mathcal{V}_{\sigma^{\mathrm{y}}}(\omega,\alpha) = 606.4$ Km/s,
% corresponding to $\gamma_{\mathrm{x}}(\omega,\alpha) = 116.3^{\circ}$
% $\gamma_{\mathrm{y}}(\omega,\alpha) = 110.1^{\circ}$ and both for
% $\alpha=150^{\circ}$, where the plots are not presented.

\subsection{Fixing the electron velocity} 
\label{sec:res-fixvel}

Here we calculated $\mathcal{V}_{\mathrm{a}}(\omega,\alpha)$ (Eq. \eqref{eq:vv-mag})
after fixing the electron velocity direction, $\mathrm{a}$, to the $x$ or $y$ direction
along the surface of the Up and Alt structures. From Eqns. \eqref{eq:polar-ang}
and \eqref{eq:azimuthal-ang}, we determined the polar, $\theta_{\mathrm{a}}
(\omega,\alpha)$, and azimuthal, $\varphi_{\mathrm{a}} (\omega,\alpha)$, angles
corresponding to the direction of the injected (?) spin.

\subsubsection{Up structure}

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{upplots/up-vx-vy-w1}
\caption{For the Up structure, we show
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$
(left scale, black line),  
$\theta_{\mathrm{a}} (\omega,\alpha)$
(right scale, red line), and
$\varphi_{\mathrm{a}} (\omega,\alpha)$,
(right scale, blue line),
vs. $\hbar\omega$, for $\alpha=35^\circ$, 
and $\mathrm{a}=x$ and $\mathrm{a}=y$.
}
\label{fig:up-vab-comp-rtp-1}
\end{figure}

For the Up structure, we find once again that  $\alpha=35^{\circ}$ maximizes the
response. In Fig. \ref{fig:up-vab-comp-rtp-1}, we plot $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and $\varphi_{\mathrm{a}}
(\omega,\alpha)$, (right scale, blue line), vs. $\hbar\omega$, for $\mathrm{a}=x,y$. 
% 
We see that for $\hbar\omega=0.084$\,eV, the response has a maximum of
$\mathcal{V}_{\mathrm{x}} (\omega,\alpha)=431.7$\,Km/s at
$\theta_{\mathrm{x}}(\omega,\alpha) = 42.5^{\circ}$, and
$\varphi_{\mathrm{x}}(\omega,\alpha) = 208.3^{\circ}$, and
$\mathcal{V}_{\mathrm{y}} (\omega,\alpha)=687.9$\,Km/s at
$\theta_{\mathrm{y}}(\omega,\alpha) = 13.9^{\circ}$, and
$\varphi_{\mathrm{y}} (\omega,\alpha) = 82.1^{\circ}$. This means that the
spin is directed upward the third Cartesian quadrant of the $xy$ plane when the
electron moves along $x$, and is almost parallel to the $xy$ plane in
the first Cartesian quadrant when it moves along $y$. 
% 
Also from this figure, we see that when the electron moves along $x$, the spin
direction is almost constant for all the energies across the peak of the
response, having
$42.5^{\circ}<\theta_{\mathrm{x}}(\omega,\alpha)<53.7^{\circ}$ and
$208.3^{\circ}<\varphi_{\mathrm{x}}(\omega,\alpha)<215.7^{\circ}$. When
the electron moves along $y$, the spin polar angle has again small variations,
$11.3^{\circ}< \theta_{\mathrm{y}}(\omega,\alpha)<13.9^{\circ}$, but the
azimuthal angle varies significantly, $82.1^{\circ}<
\varphi_{\mathrm{y}}(\omega,\alpha)<182.4^{\circ}$.

\begin{figure}[t]
\centering
\includegraphics[width=\tama]{upplots/up-vx-vy-w2}
\caption{For the Up structure we show $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and $\varphi_{\mathrm{a}}
(\omega,\alpha)$, (right scale, blue line), vs. $\hbar\omega$, for
$\alpha=35^\circ$, and $\mathrm{a}=x,y$. }
\label{fig:up-vx-vy-w2}
\end{figure}
In Fig. \ref{fig:up-vx-vy-w2}, we plot $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ vs. $\hbar\omega$, in the range where there two local maxima with opposite sign
at $\hbar\omega=1.954$\,eV and $\hbar\omega=1.957$\,eV occur.
% 
The first is the largest of the two, with
$\mathcal{V}_{\mathrm{x}} (\omega,\alpha)=61.2$\,Km/s,
$\theta_{\mathrm{x}} (\omega,\alpha)=48.3^{\circ}$, and 
$\varphi_{\mathrm{x}} (\omega,\alpha)=54.3^{\circ}$,
for the electron moving along $x$; and
$\mathcal{V}_{\mathrm{y}} (\omega,\alpha)=293.2$\,Km/s,
$\theta_{\mathrm{y}} (\omega,\alpha)=49.8^{\circ}$, and 
$\varphi_{\mathrm{y}} (\omega,\alpha)=51.9^{\circ}$
for the electron moving along $y$.
% 
For the peak at $\hbar\omega=1.957$\,eV, we obtain 
$\theta_{\mathrm{x}} (\omega,\alpha) = 129.8^{\circ}$ and 
$\varphi_{\mathrm{x}} (\omega,\alpha) = 231.7^{\circ}$, with 
$\mathcal{V}^{\mathrm{x}} (\omega,\alpha) = 54.6$\,Km/s and
$\theta_{\mathrm{y}}(\omega,\alpha) =129.3$; and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 230.7$, with 
$\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 263.7$\,Km/s.
We remark that these angles are almost
constant for all the energy values across the peak of these two local maxima, for
which the spin is directed upward in the first Cartesian quadrant of the $xy$
plane when it (who?) moves along either $x$ or $y$ directions.

\subsubsection{Alt structure}

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{altplots/alt-vx-vy-w1} % I believe that all the similar captions
% should be modified, one possible version is below
\caption{The Alt structure: the velocity 
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line), the polar angle 
$\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and the azimuthal angle
$\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue line), vs.
$\hbar\omega$, for $\alpha=150^\circ$, and $\mathrm{a}=x$ or $\mathrm{a}=y$ are shown. }
\label{fig:alt-vx-vy-w1}
\end{figure}

\begin{figure}[tb]
\centering
\includegraphics[width=\tama]{altplots/alt-vx-vy-w2}   % Change the captions accordingly the above
\caption{For the Alt structure, we show $\mathcal{V}_{\mathrm{a}}
(\omega,\alpha)$ (left scale, black line), $\theta_{\mathrm{a}}
(\omega,\alpha)$ (right scale, red line), and $\varphi_{\mathrm{a}}
(\omega,\alpha)$, (right scale, blue line), vs. $\hbar\omega$, for
$\alpha=150^\circ$, and $\mathrm{a}=x,y$. }
\label{fig:alt-vx-vy-w2}
\end{figure}

In Figs. \ref{fig:alt-vx-vy-w1} and \ref{fig:alt-vx-vy-w2}, we plot
$\mathcal{V}_{\mathrm{a}} (\omega,\alpha)$ (left scale, black line),
$\theta_{\mathrm{a}} (\omega,\alpha)$ (right scale, red line), and
$\varphi_{\mathrm{a}} (\omega,\alpha)$, (right scale, blue line), vs.
$\hbar\omega$ in two different ranges, and  for $\mathrm{a}=x,y$. In this case,
$\alpha=150^\circ$  maximizes both $\mathcal{V}_{\mathrm{x}} (\omega,\alpha)$
and $\mathcal{V}_{\mathrm{y}} (\omega,\alpha)$, as a function of $\alpha$.
% 
In Fig. \ref{fig:alt-vx-vy-w1}, the absolute maximum $\mathcal{V}_{\mathrm{x}}
(\omega,\alpha) = 301.7$\,Km/s is at
$\hbar\omega=0.720$\,eV , $\theta_{\mathrm{x}} (\omega,\alpha) =
44.5^{\circ}$ and $\varphi_{\mathrm{x}}(\omega,\alpha) = 51.2^{\circ}$;
and $\mathcal{V}_{\mathrm{y}} (\omega,\alpha) = 905.6$\,Km/s
 at $\theta_{\mathrm{y}} (\omega,\alpha) = 119.7^{\circ}$ and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 163.4^{\circ}$. 
% 
Thus, the spin is directed upward the fourth Cartesian quadrant of the $xy$
plane when the spin velocity is directed along $x$, while it is directed downward the
second Cartesian quadrant when the spin velocity is directed along $y$.
% 
Finally, in  Fig. \ref{fig:alt-vx-vy-w1}, the absolute maximum is at
$\hbar\omega=0.911$\,eV at $\mathcal{V}_{\mathrm{x}} (\omega,\alpha) =
276.3$\,Km/s, $\theta_{\mathrm{x}} (\omega,\alpha) = 154.6^{\circ}$, and
$\varphi_{\mathrm{x}}(\omega,\alpha) = 292.3^{\circ}$, and
$\mathcal{V}_{\mathrm{y}} (\omega,\alpha) = 468.6$\,Km/s at 
$\theta_{\mathrm{y}} (\omega,\alpha) = 129.2^{\circ}$, and
$\varphi_{\mathrm{y}}(\omega,\alpha) = 228.3^{\circ}$, implying that the spin
is directed downward the fourth Cartesian quadrant of the $xy$ plane when the
spin velocity is directed along $x$, while is directed downward the third Cartesian
quadrant when the spin velocity is directed along $y$.

\section{Conclusions} % (fold)
\label{sec:conclusions}

We reported the results of an \emph{ab initio} calculations for the spin velocity injection (SVI) due to the one-photon
absorption of the linearly polarized light in the Up and Alt 2D 50\% hydrogenated
graphene structures. Different possible arrangements of the of the spin injection have been considered: we made the calculations for the cases when the spin is
polarized in $z$ direction or when the velocity is directed along $x$ or
$y$. To the best of our knowledge, this effect  has been previously reported in this  2D partially hydrogenated 
structures.
% 
We have shown that the  SVI demonstrates an anisotropic behavior, which is very sensitive to the symmetry of the structures of interest. We have found that the Up structure shows
the strongest response for the spin directed along $z$, resulting in the velocity 
% 
$\mathcal{V}_{\sigma^{\mathrm{z}}} (\omega,\alpha) = 668.0$\,Km/s 
% 
for the incoming photon energy of 0.084\,eV. Also, the Alt structure
has the strongest response when the spin moves along $y$ direction,
resulting in 
% 
$\mathcal{V}^{\mathrm{y}} (\omega,\alpha) = 905.6$\,Km/s
% 
for the incoming photon energy of 0.720\,eV.
% 
The speed values obtained here are of the same order of magnitude
as those of Ref.~\onlinecite{najmaiePRB03} in 
unbiased semiconductor quantum well structures, while they are of order of magnitude 
higher compared to 3D bulk materials.
The spin relaxation time in pure and doped graphene is long enough in the order
from nanoseconds to milliseconds. \cite{wojtaszekPRB13,ertlerPRB09} Therefore,
according to our results, due to the high spin velocity transport both structures considered are the excellent candidates for the development of
spintronics devices that require pure spin current (PSC).

\section{Acknowledgment} % (fold)

This work has been supported by \emph{Consejo Nacional de Ciencia y
Tecnolog\'ia} (CONACyT), M\'exico, Grant No. 153930.
R.Z.P. thanks CONACyT for scholarship support.
\bibliography{article.bib}

\end{document}